A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-10-25 LyapDraft.tex
small changes to eat seq section
fixed all the curly braces

\begin{document}
\title{A Lyapunov Analysis of Momentum Methods \\ in  Optimization}
%\title{A Lyapunov Perspective on Momentum Methods \\ in  Optimization}
\date{}
\author{Ashia C. Wilson \and Benjamin Recht \and Michael I. Jordan}
\maketitle
\begin{abstract}
Momentum methods play a central role in convex optimization. Several momentum methods are provably optimal and all use a technique called {\em estimate sequences} to analyze their behavior.  The technique of estimate sequences has long been considered difficult to understand. In the following paper, we show there is an equivalence between the technique of estimate sequences and a family of Lyapunov functions in both continuous and discrete time. This framework allows us to develop a simplified and unified analysis of several existing algorithms and introduce a couple of new algorithms. Furthermore, it provides insights into the art of algorithm design by strengthening the connection between algorithms and continuous time dynamical systems. 
\end{abstract}

@@

Polyak motivated momentum methods by an analogy to a ``Heavy Ball'' moving in a potential well defined by the cost function.  However, Polyak's physical intuition was exceedingly difficult to make rigorous.  For quadratics costs, Polyak was able to apply an eigenvalue argument that showed that his Heavy Ball method required no more iterations than the method of conjugate gradients~\cite{Polyak1964}.  
Indeed, when applied to positive definite quadratic cost functions, Polyak's Heavy Ball method is equivalent to Chebyshev's iterative method~\cite{Chebyshev}.   Despite its intuitive elegance, Polyak's eigenvalue analysis does not apply globally for general convex functions. Indeed, Lessard~\emph{et al} derived a simple one dimensional counterexample where the standard Heavy-Ball method does not converge~\cite{Lessard14}.

@@

To overcome this lack of intuition, several authors have recently proposed schemes to achieve acceleration without appealing to estimate sequences~\cite{Fazel, BubeckLeeSingh15, Lessard14}.  Orthogonally, another set of authors have taken a different approach by analyzing the continuous-time limit of accelerated methods and showing that the stability of the resulting differential equations can be verified by analyzing a simple Lyapunov function~\cite{SuBoydCandes14,Krichene15,Acceleration}.  Unfortunately, none of these works provide a clear path of moving from a continuous time ODE to a discrete time optimization algorithm.  There are a vast number of ways to discretize differential equations, but not all of them give rise to convergent methods.  Indeed, for unconstrained optimization on Euclidean space, Polyak's Heavy Ball method and Nesterov's optimal method have the same continuous time limit.

In this paper, we propose a bridge between the continuous time limits and discrete time algorithms.  Our method takes as its primary object, a general Lyapunov function for momentum methods in continuous time.  Through a diverse set of examples, we demonstrate how the proofs of momentum methods can be understood as bounding discretization errors of the Lyapunov function when moving to discrete time.  In this way, we show how the discretization of the associated continuous time ODE needs to be performed in such a way that the Lyapunov function remains valid when transitioning to discrete time.

Using this technique, we provide a clean, direct proof of the convergence of Nesterov's method for strongly convex functions in Euclidean spaces.  We explore how only particular parameterizations of the continuous differential equation lead to discrete time methods.  In doing so, we explain the need for the extragradient step inside Nesterov's method which does not appear in Polyak's method.

Finally, we show there is an equivalence between estimate sequences and Lyapunov functions.  In continuous time and discrete time, estimate sequences can be derived directly from the Lyapunov function and vice versa.  We show that the associated continuous time estimate sequence can also be directly discretized to give standard estimate-sequence based proofs of momentum methods.  This clarifies how Nesterov's ``algebraic trick'' is closely related invariant reduction proofs that are more common in algorithm analysis.

The paper proceeds as follows.  We first describe the related work in viewing optimization algorithms as discretizations of continuous dynamical systems.  We then introduce the notion of Lyapunov functions that we will use throughout the paper to demonstrate convergence of algorithms.  In particular, we will introduce time-varying Lyapunov functions that not only prove algorithm convergence but additionally verify a \emph{convergence rate}.  We review how such Lyapunov analysis has been used to analyze mirror decent, reviewing the concepts of Bregman divergence and mirror maps in the process.  We then turn to accelerated methods, providing a general Lyapunov analysis for Nesterov's accelerated method, the quasi-monotone subgradient method, the conditional gradient method, and a few novel algorithms.  We pay particular attention to the strongly convex case in Euclidean space, highlighting some unique properties of this setup.  Finally, we describe the connection between estimate sequences and Lyapunov functions and describe directions for future investigation.

@@

The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from looking at optimization algorithms as discrete-time approximations to ordinary differential equations. In particular, the second order differential equation:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \gamma \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity as well as a potential $f$.  Note that the fixed points of such dynamics are those where the gradient of $f$ vanishes.   The trajectories of this dynamical systems will tend to continue to move in the direction they were moving before, analogous to how heavier objects move down hill faster than light objects in the presence of friction.  In the limit that the mass goes to zero, where $\gamma \propto 1/m$, we recover the continuous time limit of the gradient descent.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

In seminal work, Polyak proposed the \emph{Heavy Ball method} for optimization which amounted to applying the Euler method to the dynamical system~\eqref{eq:heavy-ball-ode}.  Polyak was able to show \emph{local convergence} of this method, but was unable to prove global convergence.  Moreover, Polyak's local convergence required the assumption that $f$ was strongly convex in the neighborhood of the associated stationary point.

@@

 Wibisono, Wilson, and Jordan~\cite{Acceleration} followed up on the work of Su, Boyd and Candes~\cite{SuBoydCandes14}, introducing a class of dynamical systems whose discretizations give rise to a family of general accelerated gradient algorithms. They showed that many accelerated methods can be viewed as a discretization of the following second-order ODE,
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,.
\end{align}
Here $h$ is a strongly convex function associated with the geometry of the space.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous time limit $m \rightarrow 0$ of this differential equation recovers the mirror descent dynamic~\cite[(78)]{Acceleration}, with $h$ playing the role of the distance generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

We briefly mention the appearance of the $e^{-\alpha_t}$ term in~\eqref{Eq:E} and~\eqref{eq:general-sc-lyap}. This unusual given most Lyapunov functions are time-invariant. In~\cite{Acceleration}, this scaling was studied at length. In particular, it was shown that the family of Lagrangian functionals which generates~\eqref{Eq:EL} is invariant under the action of time reparameterization. By contrast, the mirror descent dynamic, which is a gradient flow, is not invariant under time-reparameterization. %Under the ideal scaling \eqref{Eq:IdeSca}, time reparameterization 
This invariance property explains why the Lyapunov function remains valid for the momentum dynamic~\eqref{Eq:E} which parameterized by $\alpha_t$ and $\beta_t$. 
%in these reparameterizations of time, however, you always have a valid lyapunov function
%This invariance property might suggest why momentum family is more natural for the purposes of optimization (that is, it might help us understand why the discrete-time implementations of the family of momentum dynamics achieves systematically better rates of convergence than the family of gradient flows).

@@

\subsection{The Quasi-Monotone Subgradient Method}
\label{Sec:QuasiMono}
The quasi-monotone subgradient method was developed by Nesterov~\cite{Nesterov15} as an alternative to the dual averaging method -- though both achieve the $O(1/\sqrt{k})$ lower bound for the class of methods with bounded subgradients. The advantage of the quasi-monotone method however is that a convergence rate guarantee can be shown for the entire sequence of iterates instead of the average (or minimum) iterate, as is the case with the dual averaging method in the non-Euclidean setting. {\em We begin by studying the quasi-monotone method in the situation where we have full gradients, which we assume to be bounded}. We also set the dual averaging term $\gamma_k \equiv 1$ for convenience and give an analysis of the full algorithm in Appendix~\ref{App:ProofQuasiGen}.  With these minor modifications, the algorithm can understood as the explicit-Euler method applied to~\eqref{Eq:Z} and the implicit-Euler method applied to~\eqref{Eq:X} (where now, we take $e^{\beta_t} = A_{k+1}$):

@@

In this section, we connect our Lyapunov functions to the more standard method of estimate sequences.  This connection is formal in continuous time.  We derive continuous time estimate sequences directly from our Lyapunov function.  Moreover, we show how discretizing the estimate sequence gives a proof of convergence of the associated momentum methods.

@@

\begin{table}

@@

In this section, we show how one can show that estimate sequences are equivalent to our Lyapunov framework in both continuous and discrete time. 
The continuous time view shows that the errors in both the Lyapunov function and estimate sequences are due to discretization errors. We provide a sketch for how this works for most of these methods leaving some of the details for Appendix~\ref{App:EstSeq}.

@@

\subsection{Searching for invariants}
Earlier work by Drori and Teboulle~\cite{DroriTeboulle13}, Kim and Fessler~\cite{Kim2016}, Taylor \emph{et al}~\cite{Taylor2016}, and Lessard~\emph{et al}~\cite{Lessard14} have shown that optimization algorithms could be analyzed by solving convex programming problems.  In particular, Lessard~\emph{et al} show that Lyapunov-like potential functions called \emph{integral quadratic constraints} can be found by solving a constant-sized semidefinite programming problem. It would be interesting to see if these results can be adapted to directly search for Lyapunov functions like those studied in this paper.  This would provide a method to automate the analysis of new techniques, possibly moving beyond momentum methods to novel families of optimization techniques.

\bibliographystyle{plain}
\bibliography{refs.bib}

\appendix
 \section{Mirror Descent Dynamic}
 \label{App:Mirror}
 We begin with a smooth manifold $\X$ with a local metric $\mathrm{g}(x)$ and a function $f:\X \rightarrow \R$ we would like to minimize. The gradient flow associated with $f$ is the flow induced by the differential equation,
\begin{equation}
\dot X_t = v(X_t),
\end{equation} where the vector field $v(X_t)$ is the ``steepest descent'' direction (the direction that makes $f$ decrease the fastest)
\begin{equation}
v(x) = \arg \min _v \left\{ \langle \nabla f(X_t), v\rangle + \frac{1}{2}\|v\|_{\mathrm{g}(X_t)}^2\right\}.
\end{equation}
We can write the gradient flow equation explicitly as,
 %along which $f$ decreases the most, when measured by the metric $\mathrm{g}(X_t)$. If we unpack this statement, notice that  % which characterized the direction of steepest descent. In particular, 
%\begin{equation}
%\frac{d}{dt} f(X_t) = \langle \nabla f(X_t), \dot X_t\rangle = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), \dot X_t\rangle_x = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), v(X_t)\rangle_x.
%\end{equation} 
%From above, we can conclude that the direction $v(X_t)$, along which $f$ can be said to decrease the most when measured by the metric, is $v(X_t) = -\mathrm{g}(X_t)^{-1}\nabla f(X_t)$.  Thus, for a smooth manifold $\X$ with local metric $\mathrm{g}(x)$, the gradient flow equation can be written as
\begin{equation}\label{Eq:ManGradFlow}
\dot X_t = - \mathrm{g}(X_t)^{-1} \nabla f(X_t).
\end{equation}
%
%the interpretation of a gradient flow as a steepest descent flow becomes more obvious. 
%However, it is often thought of vector field  the metric to map the differential of the function $\nabla f(x)$ to the tangent space. 
%Intuitively, the gradient is a directional derivative and is therefore an element of the cotangent space. The metric is used to map the gradient onto the tangent space to figure out the corresponding vector field.  
%\paragraph{The Mirror Descent Dynamic} 
The dynamic which forms the basis for the mirror descent algorithm arises from a special structure that appears when the metric is chosen to be the Hessian of a strictly convex function $h$:
% $\mathrm{g}(x) = \nabla^2 h(x)$, 
\begin{equation}\label{Eq:NatGrad}
\dot X_t = -[\nabla^2 h(X_t)]^{-1} \nabla f(X_t).
\end{equation}
which we can also write as,
\begin{equation}\label{Eq:MirrorFlow}
\frac{d}{dt} \nabla h(X_t) = - \nabla f(X_t).
\end{equation}
While on the surface, the ability to rewrite \eqref{Eq:NatGrad} as \eqref{Eq:MirrorFlow} appears to be a nice algebraic trick, it is actually a manifestation of the following fact: \begin{center}{\em When $h:\X \rightarrow \R$ is strictly convex and $\mathrm{g}(x)= \nabla^2 h(x)$, there is a special map \\$\phi:\X \rightarrow \mathcal{Y}$ called the {\em mirror map}, given by $\phi = \nabla h$, whose push-forward maps\\ a gradient flow on $\X$ to a gradient flow on $\mathcal{Y}$. \cite{Acceleration}}\end{center} As a quick exercise, we can check that the following holds under the mirror map:

@@

\frac{d}{dt} \nabla h(X_t) = -\dot \tau(t)\nabla f(X_t) = - \frac{d}{dt}\left\{e^{\beta_t}\right\} \nabla f(X_t).

@@

& \overset{\eqref{Eq:MirrorTime}}{=}  \frac{d}{dt}\left\{e^{\beta_t}\right\} \left\langle \nabla f(X_t), x^\ast - X_t\right\rangle \notag\\
& \leq   \frac{d}{dt}\left\{e^{\beta_t}\right\} (f(X_t) - f(x^\ast)) dt,\label{Eq:Bound1} 

@@

\frac{d}{dt}\left\{ e^{\beta_t} f(X_t) \right\} &= \frac{d}{dt} \{e^{\beta_t}\}f(X_t) + e^{\beta_t}\langle \nabla f(X_t), \dot X_t\rangle\notag \\
&= \frac{d}{dt} \{e^{\beta_t}\}f(X_t) - \dot \beta_t e^{2\beta_t}\langle \nabla f(X_t), \nabla^2 h(X_t)^{-1}\nabla f(X_t)\rangle \notag\\
& = \frac{d}{dt} \{e^{\beta_t}\}f(X_t) - \dot \beta_t e^{2\beta_t}\frac{1}{2}\|\nabla f(X_t)\|_{\ast,X_t}^2 \notag \\
&\leq \frac{d}{dt} \{e^{\beta_t}\}f(X_t). 

@@

 \frac{d}{dt} \left\{ e^{\beta_t}( f(X_t) - f(x^\ast))\right\} \leq\frac{d}{dt} \{e^{\beta_t}\}\left( f(X_t) - f(x^\ast)\right) \leq - \frac{d}{dt} D_h(x^\ast,X_t). 

@@

@@

&= \left(1 - \frac{\alpha_k}{A_{k+1}} \right)\left(f(x_k) -\frac{\tilde \varepsilon_k}{A_k}\right) + \frac{\alpha_k}{A_{k+1}} f_k(x).

@@

f(x_{k+1}) -f(x) + \frac{\mu}{2}\|x - z_{k+1}\|^2 - \left(f(x_k) - f(x)  + \frac{\mu}{2}\|x - z_{k+1}\|^2\right) &\leq - \tau_k \left(f(x_k) - f(x)  + \frac{\mu}{2}\|x - z_{k+1}\|^2\right). 

>>>>

\begin{document}
\title{A Lyapunov Analysis of Momentum Methods \\ in  Optimization}
%\title{A Lyapunov Perspective on Momentum Methods \\ in  Optimization}
\date{}
\author{Ashia C. Wilson \and Benjamin Recht \and Michael I. Jordan}
\maketitle
\begin{abstract}
Momentum methods play a central role in convex optimization. Several momentum methods are provably optimal and all use a technique called {\em estimate sequences} to analyze their behavior.  The technique of estimate sequences has long been considered difficult to understand. In the following paper, we show there is an equivalence between the technique of estimate sequences and a family of Lyapunov functions in both continuous and discrete time. This framework allows us to develop a simplified and unified analysis of several existing algorithms, introduce a couple of new algorithms, and strengthen the connection between algorithms and continuous time dynamical systems. 
\end{abstract}

@@

Polyak motivated momentum methods by an analogy to a ``heavy ball'' moving in a potential well defined by the cost function.  However, Polyak's physical intuition was exceedingly difficult to make rigorous.  For quadratics costs, Polyak was able to apply an eigenvalue argument that showed that his Heavy Ball method required no more iterations than the method of conjugate gradients~\cite{Polyak1964}.  
Indeed, when applied to positive definite quadratic cost functions, Polyak's Heavy Ball method is equivalent to Chebyshev's iterative method~\cite{Chebyshev}.   Despite its intuitive elegance, Polyak's eigenvalue analysis does not apply globally for general convex functions. Indeed, Lessard~\emph{et al} derived a simple one dimensional counterexample where the standard Heavy-Ball method does not converge~\cite{Lessard14}.

@@

To overcome this lack of intuition, several authors have recently proposed schemes to achieve acceleration without appealing to estimate sequences~\cite{Fazel, BubeckLeeSingh15, Lessard14}.  Orthogonally, another set of authors have taken a different approach by analyzing the continuous-time limit of accelerated methods and showing that the stability of the resulting ODEs can be verified by analyzing a simple Lyapunov function~\cite{SuBoydCandes14,Krichene15,Acceleration}.  Unfortunately, none of these works provide a clear path of moving from a continuous time ODE to a discrete time optimization algorithm.  There are a vast number of ways to discretize ODEs, but not all of them give rise to convergent methods.  Indeed, for unconstrained optimization on Euclidean space, Polyak's Heavy Ball method and Nesterov's optimal method have the same continuous time limit.

In this paper, we propose a bridge between the continuous time limits and discrete time algorithms.  Our method takes as its primary object, a general Lyapunov function for momentum methods in continuous time.  Through a diverse set of examples, we demonstrate how the proofs of momentum methods can be understood as bounding discretization errors of the Lyapunov function when moving to discrete time.  In particular, we show how the discretization of the associated continuous time ODE needs to be performed in such a way that the Lyapunov function remains valid when transitioning to discrete time.

Using this technique, we provide a clean, direct proof of the convergence of Nesterov's method for strongly convex functions in Euclidean spaces.  We explore how only particular parameterizations of the continuous ODEs lead to discrete time methods.  In doing so, we explain the need for the extragradient step inside Nesterov's method which does not appear in Polyak's method.

Finally, we show there is an equivalence between estimate sequences and Lyapunov functions.  In continuous time and discrete time, estimate sequences can be derived directly from the Lyapunov function and vice versa.  We show that the associated continuous time estimate sequence can also be directly discretized to give standard estimate-sequence based proofs of momentum methods.  This clarifies how Nesterov's ``algebraic trick'' is closely related to invariant reduction proofs that are more common in algorithm analysis.

The paper proceeds as follows.  We first describe the related work in viewing optimization algorithms as discretizations of continuous dynamical systems.  We then introduce the notion of Lyapunov functions that we will use throughout the paper to demonstrate convergence of algorithms.  In particular, we will introduce time-varying Lyapunov functions that not only prove algorithm convergence but additionally verify a \emph{convergence rate}.  
%We review how such Lyapunov analysis has been used to analyze mirror decent, reviewing the concepts of Bregman divergence and mirror maps in the process. 
 We then turn to various momentum-based methods, providing a general Lyapunov analysis for Nesterov's accelerated method, the quasi-monotone subgradient method, the conditional gradient method, and a few novel algorithms.  We pay particular attention to the strongly convex case in Euclidean space, highlighting some unique properties of this setup.  Finally, we describe the connection between estimate sequences and Lyapunov functions and directions for future investigation.

@@

The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from looking at optimization algorithms as discrete-time approximations to ODEs. In particular, the second order ODE:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \gamma \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity as well as a potential $f$.  Note that the fixed points of such dynamics are those where the gradient of $f$ vanishes.   The trajectories of this dynamical systems will tend to continue to move in the direction they were moving before, analogous to how heavier objects move down hill faster than light objects in the presence of friction.  In the limit that the mass goes to zero, where $\gamma \propto 1/m$, we recover the continuous time limit of the gradient descent.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

In seminal work, Polyak proposed the Heavy Ball method for optimization which amounted to applying the Euler method to the dynamical system~\eqref{eq:heavy-ball-ode}.  Polyak was able to show \emph{local convergence} of this method, but was unable to prove global convergence.  Moreover, Polyak's local convergence required the assumption that $f$ was strongly convex in the neighborhood of the associated stationary point.

@@

 Wibisono, Wilson, and Jordan~\cite{Acceleration} followed up on the work of Su, Boyd and Candes~\cite{SuBoydCandes14}, introducing a class of dynamical systems whose discretizations give rise to a family of general accelerated gradient algorithms. They showed that many accelerated methods can be viewed as a discretization of the following second-order ODE,
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,.
\end{align}
Here $h$ is a strongly convex function associated with the geometry of the space.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous time limit $m \rightarrow 0$ of this ODE recovers the mirror descent dynamic~\cite[(78)]{Acceleration}, with $h$ playing the role of the distance generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

We briefly mention the appearance of the $e^{-\alpha_t}$ term in~\eqref{Eq:E} and~\eqref{eq:general-sc-lyap}. Given most Lyapunov functions are time-invariant, the appearance of this term in the Lyapunov function could be considered quite mysterious. In~\cite{Acceleration2}, this scaling was studied at length. In particular, it was shown that the family of Lagrangian functionals which generates~\eqref{Eq:EL} is invariant under the action of time reparameterization. By contrast, the mirror descent dynamic, which is a gradient flow, is not invariant under time-reparameterization. %Under the ideal scaling \eqref{Eq:IdeSca}, time reparameterization 
This invariance property helps to explain why the Lyapunov function remains valid for the momentum dynamic~\eqref{Eq:E} under the ideal scaling condition~\eqref{Eq:IdeSca}.  
%in these reparameterizations of time, however, you always have a valid lyapunov function
%This invariance property might suggest why momentum family is more natural for the purposes of optimization (that is, it might help us understand why the discrete-time implementations of the family of momentum dynamics achieves systematically better rates of convergence than the family of gradient flows).

@@

\subsection{The Quasi-Monotone Subgradient Method}
\label{Sec:QuasiMono}
The quasi-monotone subgradient method was developed by Nesterov~\cite{Nesterov15} as an alternative to the dual averaging method -- though both achieve the $O(1/\sqrt{k})$ lower bound for the class of methods with bounded subgradients. The advantage of the quasi-monotone method however is that a convergence rate guarantee can be shown for the entire sequence of iterates instead of the average (or minimum) iterate, as is the case with the dual averaging method in the non-Euclidean setting. {\em We begin by studying the quasi-monotone method in the situation where we have full gradients, which we assume to be bounded}. We also set the dual averaging term $\gamma_k \equiv 1$ for convenience and give an analysis of the full algorithm in Appendix~\ref{App:ProofQuasiGen}.  With these minor modifications, the algorithm can be understood as the explicit-Euler method applied to~\eqref{Eq:Z} and the implicit-Euler method applied to~\eqref{Eq:X} (where now, we take $e^{\beta_t} = A_{k+1}$):

@@

In this section, we connect our Lyapunov framework to the technique of estimate sequences.  We derive continuous time estimate sequences directly from our Lyapunov function and demonstrate how these two techniques are equivalent in discrete time.

@@

\begin{table}[h!]

@@

Now we demonstrate how these two frameworks are equivalent.  
The continuous time view shows that the errors in both the Lyapunov function and estimate sequences are due to discretization errors. We provide a sketch for how this works for most of these methods, leaving some of the details for Appendix~\ref{App:EstSeq}.

@@

\subsection{Searching for invariants}
Earlier work by Drori and Teboulle~\cite{DroriTeboulle13}, Kim and Fessler~\cite{Kim2016}, Taylor \emph{et al}~\cite{Taylor2016}, and Lessard~\emph{et al}~\cite{Lessard14} have shown that optimization algorithms could be analyzed by solving convex programming problems.  In particular, Lessard~\emph{et al} show that Lyapunov-like potential functions called \emph{integral quadratic constraints} can be found by solving a constant-sized semidefinite programming problem. It would be interesting to see if these results can be adapted to directly search for Lyapunov functions like those studied in this paper.  This would provide a method to automate the analysis of new techniques, possibly moving beyond momentum methods to novel families of optimization techniques.
\section*{Acknowledgement}
We would like to give special thanks to Andre Wibisono for the many helpful discussion involving this paper. We would also like to thank Orianna Demassi for several helpful suggestions and Stephen Tu who caught a small error in a previous version of this paper.
\bibliographystyle{plain}
\bibliography{refs.bib}

\clearpage
\appendix
 \section{Mirror Descent Dynamic}
 \label{App:Mirror}
 We begin with a smooth manifold $\X$ with a local metric $\mathrm{g}(x)$ and a function $f:\X \rightarrow \R$ we would like to minimize. The gradient flow associated with $f$ is the flow induced by the differential equation,
\begin{equation}
\dot X_t = v(X_t),
\end{equation} where the vector field $v(X_t)$ is the ``steepest descent'' direction (the direction that makes $f$ decrease the fastest)
\begin{equation}
v(x) = \arg \min _v \left\{ \langle \nabla f(X_t), v\rangle + \frac{1}{2}\|v\|_{\mathrm{g}(X_t)}^2\right\}.
\end{equation}
We can write the gradient flow equation explicitly as,
 %along which $f$ decreases the most, when measured by the metric $\mathrm{g}(X_t)$. If we unpack this statement, notice that  % which characterized the direction of steepest descent. In particular, 
%\begin{equation}
%\frac{d}{dt} f(X_t) = \langle \nabla f(X_t), \dot X_t\rangle = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), \dot X_t\rangle_x = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), v(X_t)\rangle_x.
%\end{equation} 
%From above, we can conclude that the direction $v(X_t)$, along which $f$ can be said to decrease the most when measured by the metric, is $v(X_t) = -\mathrm{g}(X_t)^{-1}\nabla f(X_t)$.  Thus, for a smooth manifold $\X$ with local metric $\mathrm{g}(x)$, the gradient flow equation can be written as
\begin{equation}\label{Eq:ManGradFlow}
\dot X_t = - \mathrm{g}(X_t)^{-1} \nabla f(X_t).
\end{equation}
%
%the interpretation of a gradient flow as a steepest descent flow becomes more obvious. 
%However, it is often thought of vector field  the metric to map the differential of the function $\nabla f(x)$ to the tangent space. 
%Intuitively, the gradient is a directional derivative and is therefore an element of the cotangent space. The metric is used to map the gradient onto the tangent space to figure out the corresponding vector field.  
%\paragraph{The Mirror Descent Dynamic} 
The dynamic which forms the basis for the mirror descent algorithm arises from a special structure that appears when the metric is chosen to be the Hessian of a strictly convex function $h$:
% $\mathrm{g}(x) = \nabla^2 h(x)$, 
\begin{equation}\label{Eq:NatGrad}
\dot X_t = -[\nabla^2 h(X_t)]^{-1} \nabla f(X_t).
\end{equation}
which we can also write as,
\begin{equation}\label{Eq:MirrorFlow}
\frac{d}{dt} \nabla h(X_t) = - \nabla f(X_t).
\end{equation}
While on the surface, the ability to rewrite \eqref{Eq:NatGrad} as \eqref{Eq:MirrorFlow} appears to be a nice algebraic trick, it is actually a manifestation of the following fact: \begin{center}{\em When $h:\X \rightarrow \R$ is strictly convex and $\mathrm{g}(x)= \nabla^2 h(x)$, there is a special map \\$\phi:\X \rightarrow \mathcal{Y}$ called the {\em mirror map}, given by $\phi = \nabla h$, whose push-forward maps\\ a gradient flow on $\X$ to a gradient flow on $\mathcal{Y}$. \cite{Acceleration}}\end{center} As a quick exercise, we can check that the following holds under the mirror map:

@@

\frac{d}{dt} \nabla h(X_t) = -\dot \tau(t)\nabla f(X_t) = - \frac{d}{dt}\left(e^{\beta_t}\right) \nabla f(X_t).

@@

& \overset{\eqref{Eq:MirrorTime}}{=}  \frac{d}{dt}\left(e^{\beta_t}\right) \left\langle \nabla f(X_t), x^\ast - X_t\right\rangle \notag\\
& \leq   \frac{d}{dt}\left(e^{\beta_t}\right) (f(X_t) - f(x^\ast)) dt,\label{Eq:Bound1} 

@@

\frac{d}{dt}\left( e^{\beta_t} f(X_t) \right) &= \frac{d}{dt} \Big(e^{\beta_t}\Big)f(X_t) + e^{\beta_t}\langle \nabla f(X_t), \dot X_t\rangle\notag \\
&= \frac{d}{dt} \Big(e^{\beta_t}\Big)f(X_t) - \dot \beta_t e^{2\beta_t}\langle \nabla f(X_t), \nabla^2 h(X_t)^{-1}\nabla f(X_t)\rangle \notag\\
& = \frac{d}{dt} \Big(e^{\beta_t}\Big)f(X_t) - \dot \beta_t e^{2\beta_t}\frac{1}{2}\|\nabla f(X_t)\|_{\ast,X_t}^2 \notag \\
&\leq \frac{d}{dt} \Big(e^{\beta_t}\Big)f(X_t). 

@@

 \frac{d}{dt} \left( e^{\beta_t}( f(X_t) - f(x^\ast))\right) \leq\frac{d}{dt} \Big(e^{\beta_t}\Big)\left( f(X_t) - f(x^\ast)\right) \leq - \frac{d}{dt} D_h(x^\ast,X_t). 

@@

Therefore, for~\eqref{Eq:Lyap3} we can ensure the following

@@

&\overset{\text{Tab}~\ref{table:Table}}{=} \left(1 - \frac{\alpha_k}{A_{k+1}} \right)\left(f(x_k) -\frac{\tilde \varepsilon_k}{A_k}\right) + \frac{\alpha_k}{A_{k+1}} f_k(x).

@@

f(x_{k+1}) -f(x) + \frac{\mu}{2}\|x - z_{k+1}\|^2 - \left(f(x_k) - f(x)  + \frac{\mu}{2}\|x - z_{k+1}\|^2\right) &\overset{\text{Tab}~\ref{table:Table}}{\leq} - \tau_k \left(f(x_k) - f(x)  + \frac{\mu}{2}\|x - z_{k+1}\|^2\right). 
